---
title: "A Quarto Template for arXiv Preprints"
subtitle: "Rendering to Typst without LaTeX"
date: today
date-format: long
author:
  - name: Ada Lovelace
    email: ada@example.org
    equal-contributor: true
    affiliations:
      - id: analytical
        name: Analytical Engine Institute
        department: Department of Computation
        city: London
        country: United Kingdom
  - name: Kurt Gödel
    email: kurt@example.org
    equal-contributor: true
    affiliations:
      - id: ias
        name: Institute for Advanced Study
        city: Princeton
        country: United States
  - name: Emmy Noether
    email: emmy@example.org
    affiliations:
      - id: ias
  - name: Stephen D. Turner
    email: stephen@example.org
    orcid: 0000-0001-9140-9028
    corresponding: true
    affiliations:
    - id: ias
    - id: uva
      name: University of Virginia
      department: School of Data Science
      city: Charlottesville
      country: United States
abstract: "The abstract paragraph should be indented ½ inch on both the left- and right-hand margins. Use 10 point type, with a vertical spacing (leading) of 11 points. The word Abstract must be centered, bold, and in point size 12. Two line spaces precede the abstract. The abstract must be limited to one paragraph. This document demonstrates the `arxiv-typst` Quarto format: sections, math, figures, tables, executed R and Python code, cross-references, citations, footnotes, and an appendix."
keywords: [machine learning, typesetting, reproducibility]
bibliography: references.bib
notice: "Preprint. Under review since 1987."
hide-emails: false
lineno: false
table-stripes: true
format:
  arxiv-typst:
    echo: true
    papersize: us-letter
    margin:
      x: 1in
      y: 1in
    section-numbering: "1.1"
    toc: false
    toc-title: Contents
    keep-typ: false
---

# Introduction {#sec-intro}

Deep learning has seen wide adoption since the introduction of convolutional architectures [@krizhevsky2012imagenet] and attention mechanisms [@vaswani2017attention]. This document is not a real paper; it exists to exercise the formatting features of the `arxiv-typst` format so you can check that the output matches what you expect from an arXiv preprint. The method appears in @sec-method, computational results in @sec-experiments, and supplementary material in @sec-supplement.

## Citations

Citations render in bracketed numeric style by default, e.g., generative adversarial networks [@goodfellow2014gan] or multiple works at once [@krizhevsky2012imagenet; @vaswani2017attention]. Use `@vaswani2017attention` for an in-text citation: @vaswani2017attention showed that attention suffices.

# Method {#sec-method}

Given inputs $x_1, \ldots, x_n$, we compute attention weights by a softmax over scaled dot products:

$$
\mathrm{Attention}(Q, K, V) = \mathrm{softmax}\left(\frac{QK^\top}{\sqrt{d_k}}\right)V
$$ {#eq-attn}

As shown in @eq-attn, the scaling factor $\sqrt{d_k}$ prevents the dot products from growing large in magnitude.

## Static figures {#sec-figures}

All artwork must be neat, clean, and legible. @fig-loss shows a placeholder training curve included as an image file, referenced with `@fig-loss`.

![Sample figure caption. Training loss decreases over time, as training losses tend to do.](loss-curve.svg){#fig-loss width="50%"}

## Static tables {#sec-tables}

Tables should have a caption above them, as in @tbl-results, which is a plain markdown pipe table labeled by putting `{#tbl-results}` on its caption line and referenced with `@tbl-results`. Header rows render bold automatically, and setting `table-stripes: true` in the document metadata shades alternating body rows. Avoid vertical rules.

| Model       | Params (M) | Accuracy (%) |
|:------------|-----------:|-------------:|
| Baseline    |         12 |         71.3 |
| Ours (small)|         13 |         74.8 |
| Ours (large)|         88 |         79.2 |

: Test accuracy on the held-out set. Best result in each column is not bolded because this is an example. {#tbl-results}

# Computational results {#sec-experiments}

This section holds executed code. Everything here is computed at render time by knitr, so @fig-cars, @tbl-kable, and @tbl-gt always reflect the code shown. Label a chunk `fig-*` or `tbl-*` and give it a `fig-cap` or `tbl-cap` to make it cross-referenceable.^[Footnotes appear at the bottom of the page with a short separator rule, as in the LaTeX style.]

## A figure from base R

@fig-cars plots the built-in `cars` data with a least-squares fit, using nothing but base graphics. Here we're using `echo: false` to suppress printing of the code used to generate this section.

```{r}
#| label: fig-cars
#| fig-cap: "Stopping distance against speed for the built-in cars data, with a least-squares fit."
#| fig-width: 5
#| fig-height: 2.5
#| out-width: "70%"
#| echo: false
par(mar = c(4, 4, 1, 1))
plot(
  cars,
  pch = 19,
  col = "steelblue",
  xlab = "Speed (mph)",
  ylab = "Stopping distance (ft)"
)
abline(lm(dist ~ speed, data = cars), lwd = 2, col = "firebrick")
```

## A table from knitr::kable

@tbl-kable summarizes the same data by speed quartile. `knitr::kable()` emits a markdown table, which the format styles like any other table.

```{r}
#| label: tbl-kable
#| tbl-cap: "Stopping distance by speed quartile of the cars data, via knitr::kable."
q <- cut(
  cars$speed,
  quantile(cars$speed),
  include.lowest = TRUE,
  labels = c("Q1", "Q2", "Q3", "Q4")
)
s <- aggregate(dist ~ q, data = cbind(cars, q), function(x) {
  c(mean = mean(x), sd = sd(x))
})
out <- data.frame(
  Quartile = s$q,
  Mean = round(s$dist[, "mean"], 1),
  SD = round(s$dist[, "sd"], 1)
)
knitr::kable(out)
```

## A table from gt

@tbl-gt renders the head of `mtcars` through the gt package. The `tab_options()` call keeps the render quiet: `table.font.names` replaces gt's default web font stack (Typst warns about fonts it can't find), and `stub.font.weight` replaces the default CSS keyword `initial`, which Quarto's Typst writer warns about on every stub row.

```{r}
#| label: tbl-gt
#| tbl-cap: "First rows of mtcars, via the gt package."
gt::gt(head(mtcars[, 1:5]), rownames_to_stub = TRUE) |>
  gt::tab_options(
    table.font.names = "TeX Gyre Termes",
    stub.font.weight = "normal"
  )
```

## A Python chunk

Python chunks work in the same document (here run with `python.reticulate: false`, so only the standard library is required).

```{python}
#| label: py-stats
#| python.reticulate: false
import statistics
speeds = [4, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 22, 24, 25]
print(f"mean={statistics.mean(speeds):.2f} stdev={statistics.stdev(speeds):.2f}")
```

# Discussion

Lists work as expected:

- The margins are 1 inch on all sides.
- The body font is 10 pt Times (or an OpenType lookalike).
- The first-page footer carries the preprint or conference notice.

Block quotes are indented on both sides:

> The paper is limited to nine pages, including figures. Additional pages containing references and appendices do not count toward the limit.

# Conclusion

Rendering this document should produce a PDF that looks like a classic arXiv preprint. Set `notice` in the document metadata to put any text in the first-page footer, or `anonymous: true` for an anonymized submission with line numbers.

The empty `#refs` div below places the reference list here, before the appendix, instead of at the end of the document.

::: {#refs}
:::

::: {.content-visible when-format="typst"}
```{=typst}
#pagebreak(weak: true)
#show: appendix
```
:::

# Supplementary Material {#sec-supplement}

Anything after the `#show: appendix` raw block is numbered as an appendix, and the explicit `#pagebreak(weak: true)` above it starts it on a new page (drop that line to let the appendix run on). Cross-references keep working: this is @sec-supplement, and @sec-intro still points back to the introduction.

## Additional Results

Appendix subsections are numbered A.1, A.2, and so on.
